### 2010-04-30 paper/paper.tex
Starting in on the paper

\documentclass{article}

\usepackage{url}

\title{Drawing Trees and Animating Tree Changes}

\author{Sandro Badame, Peter Boothe$^*$\\
 Manhattan College\\
\url{{sbadame.student,peter.boothe}@manhattan.edu}}

\begin{document}
\maketitle

\begin{abstract}
summary.  poster verbage
\end{abstract}

\end{document}